Do return series have power-law tails with the same index?

### 2017-04-26 papers/FX/draft.tex
Added Thomas' changes to the manuscript.

more prone to negative than to positive news in te market. 

@@

data from 1 January 2010 to {\red why 1 January?} 1 January 2015.

@@

estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). It turns out 

@@

for daily or, more generally, low \fre y 

@@

GARCH models as the CCC or BEKK ensure that the marginal \ds s have power-law tails with the same index.

@@

(such as return series in a given sector of the S\&P 500 index) may have tail indices which are close to each other.

@@

the order statistices $X_{(1)}\le \cdots\le X_{(n)}$  and construct the

@@

    {1 \over k} \sum_{i=1}^k \ln {X_{(n-i+1)} \over X_{(n-k)}}

@@

order statistics. As  matter of fact,

@@

    from 2010-01-01 to 2015-01-01 and comprise $n=1304$ observations.

@@

    Each circle corresponds to a Hill estimate $\hat\alpha_k$; the grey

@@

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\beao
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha} X_{(n-k)}\,.
\eeao
{\blue
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, theorem 2.2.1), one can show
\begin{eqnarray*}
  \sqrt k (\hat K_k - K) &\overset{d}{\to}& N\left(
    0, \left( {K \over \alpha}  \right)^2
  \right) \\
  \sqrt k (\hat K_k^\alpha - K^\alpha) &\overset{d}{\to}& N\left(
    0, \left( {K^{2\alpha}}  \right)^2
  \right)
\end{eqnarray*}
where the tail index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of the estimate of $K$
and $K^\alpha$ can be constructed. 
Estimates of $K$ of the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
\begin{figure}[htb!]
  \centering
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_K.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_K.pdf}
  \end{minipage}
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Consumer_Staples_scale.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.33\linewidth}
    \includegraphics[width=\textwidth]
    {Information_Technology_scale.pdf}
  \end{minipage}
  \caption{\small Estimates of $K$ and $K^\alpha$ of stocks in sectors
    of the S\&P 500 index. The points are the estimated values; the bars are
    the $2\sigma$ confidence intervals. The confidence bands of the
    corresponding Hill estimates $\hat \alpha$ of these sectors are
    shown in figure~\ref{fig:1}. Note $\log_{10}(K^\alpha)$ is shown
    instead of $K^\alpha$.
  }
  \label{fig:sectors_parameters}
\end{figure}
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd column of
figure~\ref{fig:sectors_parameters}.

Secondly, one can see that the values of $\alpha$ and $K$ appear
positively correlated. This prompts Pareto tails of the return
distributions. We expain more about this in
section~\ref{sec:pareto_tail}.

Thirdly, figure~\ref{fig:sectors_parameters} shows the values of $K$
in the ``Energy'' and ``Information Technology'' sectors are on
average larger than those in the ``Consumer Staples'' sector. A larger
value of $K$ means, for a given value of $p$, i.e. a given probability
of losses, large losses are more probable. Thus one can conclude that
these two sectors are considerablly riskier than the ``Consumer
Staples'' sector. This is of course a confirmation of one's economic
instinct.

Yet another indication from figure \ref{fig:sectors_parameters} is
that, while the ``Energy'' and the ``Information Technology'' sectors
are similar in riskiness, the correlation between $\alpha$ and $K$ is
stronger in ``Energy''. As discussed in the last section, moving along
a curve of equal preference in the direction of increasing $\alpha$, $K$ also
increases. So the strong positive correlation seen in the ``Energy''
sector suggests these stocks might have very similar investor
preferences. This in turn may be attributed to stronger business
relations between the energy enterprises. While two IT companies
may provide a variety of products and services and not depend on each
other, two energy companies are more likely to depend on each other
via relations of supplier and customer or otherwise to compete with each
other if they are on the same link of the chain of energy production
and distribution.

\subsection{A test of equal tail indices based on Hill estimator}
Suppose we have two weakly dependent series $X_1, \dots, X_n \sim F_X$ and
$Y_1, \dots, Y_n \sim F_Y$. Both $F_X$ and $F_Y$ have power-law tails
with indices $\alpha_X$ and $\alpha_Y$, respectively. Assume also
$X_i$ is independent of $Y_j$ for all $1 \leq i \leq j \leq n$.

@@

where $\hat \alpha_X$ and $\hat \alpha_Y$ are Hill estimates of
$\alpha_X$ and $\alpha_Y$. Then it follows from continuous mapping

@@

  Z_X - Z_Y \sim N(0, \alpha_X^2 + \alpha_Y^2)

@@

The weak convergence of \eqref{eq:x1} allows one to construct a test
of the equality $\alpha_X = \alpha_Y$. As the null hypothesis, assume
$\alpha_X = \alpha_Y$. Under this hypothesis, \eqref{eq:x1} means the
test statistic $\hat \alpha_X - \hat \alpha_Y \overset{d}{\to} N(0,
\alpha_X^2 + \alpha_Y^2)$. 

@@

``Consumder Staples'' and ``Information Technology''
sectors of the S\&P 500 index. The results are shown in
\ref{fig:PairTest}.

@@

  \caption{
    {\em 1st row}: Test for changing tail index using Hill estimator based on
    concatenated series in the ``Energy'' and ``Consumer Staples''
    sectors of S\&P 500. The green, blue and red colors represent,
    respectively, the cases when the 85\%, 90\% and 95\% quantiles of
    the limit distribution of the test statistic is exceeded. 
    Grey points stand for pairs for which the test statistic is below
    the 85\% \asy\ quantile;
    {\em 2nd row:} Test for changing tail index or scale parameter
    using Hoga's test based on the same concatenated series. black points represent

@@

    The same number of upper-order statistics are used for both tests.

@@

One can see from the figure that tail indices of equities in the
``Energy'' or ``Information Technology'' sectors vary more than do
their counterparts in the ``Consumer Staples'' sector, as the null
hypothesis is rejected more often for members of these two sectors.
Moreover, these figures suggest that the test based on Hill estimator
is more powerful in distinguishing differences between tail indices,
as it results in more rejections for the ``Energy'' and the
``Information Technology'' sectors than does the other test. We
argue this lack of power of the latter can be the result of slow
convergence of the distribution of its test statistic to the
asymptotic distribution. We elaborate on this in section~\ref{sec:Hoga}.

Nevertheless, one should bear in mind that the weak convergence
\eqref{eq:x3} is valid only on condition that both the $X_i$ and $Y_i$
series are independent of each other, which is generally untrue for
two return series in the same market.
}

\subsection{Tests for a change in the extreme tail}\label{sec:Hoga}
To the best of our knowledge, there do not exist published statistical
tests for comparing the tail indices or scale parameters of two or
more dependent \ts\ in a direct way. In addition to the test described
in the previous section, we borrow a test from

@@

To check whether any two series share the same tail index and scale parameter we concatenate

@@

``Consumer Staples'' sector we summarize the results in Figure~\ref{fig:PairTest}.

@@

\par
To get an idea about the power of the test we run it on a
sample concatenated from two independent iid samples of the same size
$n=1304$ as the S\&P 500 series. Both pieces are $t$-distributed with
distinct degrees of freedom. The results are shown in
Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the
larger the minimum tail index in the concatenated pair.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{t_sim_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
  \end{minipage}
  \caption{
    {\em Left}: Test  of concatenated $t$-samples with different
    degrees of freedom. Numbers on the axes are the degrees of freedom
    in the subsamples. For an interpretation of the colored bullets,
    see Figure~\ref{fig:PairTest}. The graph shows the limited power
    of the test. In particular, if both degrees of freedom are
    relatively large it loses the capablity of distinguishing the \ds\ .
    {\em Right}: Comparison of the \asy\ distribution of the test statistic
    $T_n$ in \eqref{eq:4} under the null hypothesis and the \ds\ of
    $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees
    of freedom.
  }
  \label{fig:t_sim_pair}
\end{figure}

@@

for an iid \seq\ of pairs $(A_t,B_t)$, $t\in\bbz$, with non-negative components satisfying $\E [A_1^\alpha]=1$.  
In the model \eqref{eq:5} we can choose $Y_t=\sigma_t^2$, $A_t=\alpha_1\,Z_{t-1}^2$ and $B_t=\alpha_0$ to achieve
\eqref{eq:6}. In turn, by an application of Breiman's lemma (see \cite{buraczewski:damek:mikosch:2016})

@@

with non-negative volatility $\sigma_{i,t}$ and an iid bivariate noise \seq\ $(\bfZ_t)$ with zero mean and unite variance components.

@@

are not natural since they do not allow for a linear relationship between all squared volatilities

@@

\subsection{An utility based argument for equal tail indices}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C < \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$.
{\blue 
 Here $v$ can be thought of as the certainty payoff equivalent to the
 risky payoff $C$; $\delta$ tunes the threshold of {\em
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappoinment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
}
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:u}
u(x)=\frac{1}{1 - \gamma}\,x^{1 - \gamma}\,,\quad \gamma>0\,. 
\eeam
In practice it is more often to observe $\gamma > 1$ {\blue Let's
consult Casper for references.} Therefore we adopt a more convenient
notation in this case
\beao
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>0\,. 
\eeao

For the sake of argument, we assume that an investor initially has one
unit of wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
bond with interest rate $r>0$ and $\phi$ units in a risky asset with
return $X$ over one time unit, i.e.,
\beam\label{eq:xxie1}
  C = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
\eeam
Then we have 
\beao
\wt u(\phi) &=& \E [u(C)] + b\, \E \big[u(C)\1(C < \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
\eeao
where $F_X$ is the \df\ of $X$ with density $f_X$, and
\beao
  q = %q(r, \phi) &:=& \ln \left( {
      %\delta v - (1 - \phi) e^r
      %\over
      %\phi
    % \right) 
\log\Big(
    \ex^r + {\delta \,v - \ex^r \over \phi}
  \Big)\,.
\eeao
Note that $C < \delta v$ \fif\ $X < q$. 
\par
Naturally, if an agent invests in
a risky asset instead of a riskless bond, he expects to obtain a
higher (on average) return from the risky asset than he is guaranteed from the
riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
$q > r $. For given $b,\delta,v$, $\wt u$ is a functional of $\phi$ and $F_X$,
$
\wt u=\wt u (F_X, \phi)\,,
$
and we assume that
\beao
\wt u_{\rm max}=\wt u_{\rm max}(F_X) = \max_{0 \le  \phi \leq 1} \wt u (F_X,\phi)
\eeao
exists.

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where 
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(F_X,\phi)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x < q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
{\blue
\begin{remark}\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pp} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\end{remark}
}
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We illustrate the results of remark~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
\fct\ \eqref{eq:u}.
For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
quantity is denoted by $I(\alpha, K)$.
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure \ref{fig:sectors_parameters}.
\par
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
This is shown in 
Figure \ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
\begin{figure}
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
  \caption{The graphs show $\hat\phi$, the optimal equity allocation
    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
    the utility $u$ from \eqref{eq:u}
    function $u(x) = -x^{-\xi}/\xi$ for $\xi = 1/2$ (left) and $\xi = 4$
    (right).
  }
  \label{fig:phi_hat_pareto}
\end{figure}

>>>>

more prone to negative than to positive news in the market. 

@@

data from 1 January 2010 to 31 December 2014.

@@

estimation of the scale parameters in the tail (these are the constants $c_+$ and $c_-$ in \eqref{eq:1}). {\red check} It turns out 

@@

for daily or, more generally, low-\fre y 

@@

GARCH models as the CCC ensure that the component-wise marginal \ds s have power-law tails with the same index.

@@

(such as return series in a given sector of the S\&P 500 index) may have tail-indices which are close to each other.

@@

the order statistics $X_{(1)}\le \cdots\le X_{(n)}$  and construct the

@@

    {1 \over k} \sum_{i=1}^k \log {X_{(n-i+1)} \over X_{(n-k)}}

@@

order statistics. As  a matter of fact,

@@

    from 1 January 2010 to 31 December 2011 and comprise $n=1304$ observations.

@@

    Each circle corresponds to a Hill estimate $\hat\alpha_{50}$; the grey

@@

\subsection{Hill estimates of lower-tail scale parameters}\label{sec:HillScaleEstimates}
We assume the pure Pareto model \eqref{eq:3} with scale parameter $K>0$.
Hill \cite{hill1975simple} proposed  the maximum-likelihood estimator of $K$
derived from the joint \ds\ of $k$ upper order statistics in the sample; cf.
Embrechts et al. \cite{embrechts:klueppelberg:mikosch:1997}, p.~334.
It is given by
\hat K_k = \left({k \over n}\right)^{1/\hat\alpha_k} X_{(n-k)}\,.
Using the asymptotic normality property of upper order statistics
(cf. de Haan and Ferreira~\cite{haan:ferreira:2006}, Theorem 2.2.1), one can show
  \sqrt{k}\, (\hat K_k - K) &\std& N\big(
    0, (K /\alpha)^2\big)\quad\mbox{and}\quad
  \sqrt {k} \,(\hat K_k^\alpha - K^\alpha) \overset{d}{\to} N\big(
    0, K^{2\alpha}  \big)\,,
\end{eqnarray*}
where the tail-index $\alpha$ is regarded as known. From the above
asymptotic normality property, confidence bands of $\hat K_k$
and $\hat K_k^{\hat\alpha}$ can be constructed.   
Estimates $\hat K_k$ in the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index are computed
using this method. The results are shown in
Figure~\ref{fig:sectors_parameters}.
  \centering
    \includegraphics[width=\textwidth]
    {Energy_K.pdf}
    \includegraphics[width=\textwidth]
    {CS_K.pdf}
    \includegraphics[width=\textwidth]
    {IT_K.pdf}
    \includegraphics[width=\textwidth]
    {Energy_scale.pdf}
    \includegraphics[width=\textwidth]
    {CS_scale.pdf}
    \includegraphics[width=\textwidth]
    {IT_scale.pdf}
  \caption{\small Estimates $\hat K_k$ and $\hat K_k^{\hat \alpha}$ on $\log_{10}$-scale 
of stocks in sectors
    of the S\&P 500 index. The points are the estimated values, the bars 
    the \asy\ 95\%-confidence intervals; the confidence bands of the
    corresponding Hill estimates $\hat \alpha_k$ of these sectors are
    shown in Figure~\ref{fig:1}. 
  \label{fig:sectors_parameters}
One can see that $K$ generally takes a rather small value. For the
more volatile sectors of ``Energy'' and ``Information Technology'',
the average value of $K$ is around 0.01, while for the more stable
sector of ``Consumer Staples'', the average value is around 0.005. Due
to the smallness of $K$, mild variations of $\alpha$ would lead to
huge variations of $K^\alpha$, as shown in the 2nd row of
Figure~\ref{fig:sectors_parameters}.
Secondly, it appears that there is positive dependence between the  values of $\alpha$ and $K$. 
This prompts Pareto tails of the return
distributions. For more explanations, see
Section~\ref{sec:pareto_tail}.
Thirdly, Figure~\ref{fig:sectors_parameters} shows that, on
average, the values of $K$
in the ``Energy'' and ``Information Technology'' sectors are  larger than those in the ``Consumer Staples'' sector. 
For a given loss \pro y, 
a larger
value of $K$ implies that large losses are more probable. Thus one can conclude that
these two sectors are considerably riskier than the ``Consumer
Staples'' sector. {\red Let Casper decide about this sentence.} This is of course a confirmation of one's economic
instinct.
Yet another indication from Figure \ref{fig:sectors_parameters} is
that, while the ``Energy'' and the ``Information Technology'' sectors
are similar in riskiness, the dependence  between $\alpha$ and $K$ is
stronger in ``Energy''. As discussed in Section~\ref{sec:3} below, moving along
a curve of equal preference in the direction of increasing $\alpha$, $K$ also
increases. So the strong positive correlation seen in the ``Energy''
sector suggests these stocks might have very similar investor
preferences. This in turn may be attributed to stronger business
relations between the energy enterprises. While two IT companies
may provide a variety of products and services and not depend on each
other, two energy companies are more likely to depend on each other
via relations of supplier and customer or otherwise to compete with each
other if they are on the same link of the chain of energy production
and distribution.

\subsection{A test for equal tail-indices based on Hill estimation}
Suppose we have two independent strictly stationary positive series $X_1, \dots, X_n $ and
$Y_1, \dots, Y_n$ with corresponding \df s $F_X$ and $F_Y$ that have power-law tails
with indices $\alpha_X$ and $\alpha_Y$, respectively. 

@@

where $\hat \alpha_X$ and $\hat \alpha_Y$ are Hill estimators of
$\alpha_X$ and $\alpha_Y$; we suppress their dependence on $k$. Then it follows from the continuous mapping

@@

  Z_X - Z_Y \sim N(0, \alpha_X^2 + \alpha_Y^2)\,.

@@

This relation allows one to construct an \asy\ test under 
the null hypothesis $\alpha_X = \alpha_Y$ and with test statistic $\hat \alpha_X - \hat \alpha_Y$.

@@

``Consumer Staples'' and ``Information Technology''
sectors of the S\&P 500 index. The results are shown in the top row of
Figure~\ref{fig:PairTest}.

@@

  \caption{{\em Top row}: Test for pairwise equality of tail-indices in the 
    ``Energy'', ``Consumer Staples'' and ``Information Technology'' 
    sectors of S\&P 500. The test statistic in \eqref{eq:x1} is based on Hill estimates 
    of $\alpha_X$ and $\alpha_Y$. 
    The green, blue and red points correspond to pairs of stock in a sector
    when the test statistic is outside the intervals $[q_{0.075},q_{0.925}]$,
 $[q_{0.05},q_{0.95}]$,  $[q_{0.025},q_{0.975}]$, repectively, where $q_p$ is the $p$-quantile of the limiting
$N(0,\alpha_X^2+\alpha_Y^2)$-\ds\ of the test statistic in \eqref{eq:x1}. 
    Grey points stand for pairs for which the test statistic is inside $[q_{0.075},q_{0.925}]$. 
    {\em Bottom row:} Test for changing tail index or scale parameter     
    using Hoga's test based on concatenated series of pairs of stocks. The green, blue and red points 
correspond to pairs of stock in a sector 
    when the test statistic $T_n$ is exceeds the $85\%$-, $90\%$-, $95\%$-quantile of the limit \ds .  
  Grey points stand for pairs for which the test statistic is below the \asy\ $85\%$-quantile. Black points represent

@@

    The same number of upper order statistics are used for both tests.}

@@

It indicates that tail indices of equities in the
``Energy'' or ``Information Technology'' sectors are more variable 
in the ``Consumer Staples'' sector, as the null
hypothesis is rejected more often for members of these two former sectors.
Moreover, these figures suggest that the test based on the Hill estimator
is quite powerful in distinguishing between tail indices. In contrast to the test presented in Section~\ref{sec:Hoga}
the present test results in  more rejections for the ``Energy'' and the
``Information Technology'' sectors.
As a caution, one should bear in mind that 
\eqref{eq:x3} is valid on condition that the $X$- and $Y$-
series are independent of each other (or weakly dependent on each other), which is generally untrue for
two return series in the same market.

\subsection{A tests for a change in the extreme tail}\label{sec:Hoga}
Here we apply a test from

@@

To check whether any pair of stocks shares the same tail index and scale parameter we concatenate

@@

``Consumer Staples'' sector we summarize the results in the bottom row of Figure~\ref{fig:PairTest}.

@@

\par
To get an idea about the power of the test we run it on a
sample concatenated from two independent iid samples of the same size
$n=1304$ as the S\&P 500 series. Both pieces are $t$-distributed with
distinct degrees of freedom. The results are shown in
Figure~\ref{fig:t_sim_pair}: the power of the test is the smaller the
larger the minimum tail index in the concatenated pair.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{t_sim_pair.pdf}
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{Hoga_AsymptoticDistribution.pdf}
  \end{minipage}
  \caption{
    {\em Left}: Test  of concatenated $t$-samples with different
    degrees of freedom. Numbers on the axes are the degrees of freedom
    in the subsamples. For an interpretation of the colored bullets,
    see the bottom row of Figure~\ref{fig:PairTest}. The graph shows the limited power
    of the test. In particular, if both degrees of freedom are
    relatively large it loses the capablity of distinguishing between the \ds s.
    {\em Right}: Comparison of the \asy\ distribution of the test statistic
    $T_n$ in \eqref{eq:4} under the null hypothesis and the \ds\ of
    $T_n$ for $n=1304$ iid $t$-distributed $X_t$ with 3 and 4 degrees
    of freedom.
  }
  \label{fig:t_sim_pair}
\end{figure}

@@

for an iid \seq\ of pairs $(A_t,B_t)$, $t\in\bbz$, with non-negative components satisfying $\E [A_1^{\alpha/2}]=1$.  
In the model \eqref{eq:5} we can choose $Y_t=\sigma_t^2$, $A_t=\alpha_1\,Z_{t-1}^2+\beta_1$ and $B_t=\alpha_0$ to achieve
\eqref{eq:6}. In turn, by an application of Breiman's lemma (see \cite{buraczewski:damek:mikosch:2016}, p.~275)

@@

with non-negative volatility $\sigma_{i,t}$ and an iid bivariate noise \seq\ $(\bfZ_t)$ with zero mean and unit variance components.

@@

are degenerate since they do not allow for a linear relationship between all squared volatilities

@@

\subsection{An utility based argument for equal tail indices}\label{sec:3}
In this section we give an argument based on economic theory which  favors equality of  tail indices for return series.
We follow an approach by Routledge and Zin \cite{routledge2010generalized} who introduced the notion of
Generalized Disappointment Aversion (GDA). We consider the risky payoff $C$ of an investor and assume that
it has a continuous \ds\ on $(0,\infty)$ with \df\ $F_C$. Let $u$ be a utility \fct\ assumed to be increasing and concave on $(0,\infty)$.
Following Routledge and Zin \cite{routledge2010generalized},
the utility of an agent with GDA preferences is given by
\beao% \label{eq:xxie0}
  \wt u&=& \E [u(C)] - b\, \int_{0}^{\delta v}
  \big[ u(\delta \,v) - u(x) \big] F_C(dx)\nonumber\\&=& \E u(C) - b \E\big[\big(u(\delta \,v) - u(C)\big)\1( C \le  \delta \,v)\big]\,,
\eeao
where $\delta $ and $v$ are positive constants, and $b\ge 0$.
  Here $v$ can be thought of as the {\em certainty payoff} equivalent to the
 risky payoff $C$; $\delta$ tunes the {\em threshold of 
   disappointment} in proportion to $v$; $b$ determines the extra
 weight given to the expected return of $C$ when $C$ is below the
 disappoinment threshold $\delta v$.
 If $b=0$, preferences are the classical expected utility. If
 $\delta=1$ and $b>0$ preferences follow Gul's \cite{gul:1991} 
disappointment aversion which were generalized by Routledge and Zin
\cite{routledge2010generalized}.
\par
An agent guided by the utility function $u$ will seek to maximize the
\fct al $\wt u$.
Routledge and Zin assumed a power-law utility function 
\beam\label{eq:u}
u(x)=-\frac{1}{\xi}\,x^{-\xi}\,,\quad \xi>-1\,. 
\eeam
%In practice it is more often to observe $\gamma > 1$ {\blue Let's
%consult Casper for references.} Therefore we adopt a more convenient
For the sake of argument, we assume that an investor initially has one
unit of wealth. He invests $1-\phi\in (0,1)$ units in a risk-free 
bond with interest rate $r>0$ and $\phi$ units in a risky asset with
return $X$ over one time unit, i.e.,
\beam\label{eq:xxie1}
  C = (1 - \phi)\, \ex^{r} + \phi \,\ex^{X}\,.
\eeam
Then we have 
\beao
\wt u(\phi,F_X) &=& \E [u(C)] + b\, \E \big[u(C)\1(C \le  \delta v)\big] - b \,u(\delta\, v) F_X(q)\,, %\label{eq:functional}
\eeao
where $F_X$ is the \df\ of $X$ with density $f_X$, and
\beao
  q = %q(r, \phi) &:=& \log \left( {
      %\delta v - (1 - \phi) e^r
      %\over
      %\phi
    % \right) 
\log\Big(
    \ex^r + {\delta \,v - \ex^r \over \phi}
  \Big)\,.
\eeao
Note that $C \le \delta v$ \fif\ $X \le  q$. 
\par
Naturally, if an agent invests in
a risky asset instead of a riskless bond, he expects to obtain a
higher (on average) return from the risky asset than he is guaranteed from the
riskless bond. In our notation, this means $\delta \,v > \ex^r$ or
$q > r $. For given $b,\delta,v$, $\wt u$ is only a functional of $\phi$ and $F_X$,
$\wt u=\wt u (F_X, \phi)$.
We assume that
\beao
\wt u_{\rm max}=\wt u_{\rm max}(F_X) = \max_{0 \le  \phi \leq 1} \wt u (F_X,\phi)
\eeao
exists.

\subsection{Pareto-distributed returns}\label{sec:pareto_tail}
Since we are interested in the influence of heavy-tailed losses on the preferences of an investor 
we assume the following toy model. 
We consider the case when $X$ has a two-sided Pareto \ds\ given by
\beam\label{eq:pareto}
  F_X(x) = \left\{
  \begin{array}{ll}
    p \left(
    {K \over K - x}
    \right)^\alpha & x \leq 0 \,,\\
    1 - (1 - p) \left(
    {K' \over K' + x}
    \right)^\beta & x > 0\,,
  \end{array}
  \right.
\eeam
where {\red where needed?}
$\alpha, \beta > 1$, $0 < p < 1$ and $K,K' > 0$. Then we have
\beam\label{eq:xxie1.0}
 \wt u(\phi,F_X)\nonumber
  &=&
  \alpha\, K^\alpha\,  p\,
  \int_{-\infty}^0
  u\big( (1 - \phi) \ex^r + \phi \ex^x \big)\,
  {1 + b \over (K - x)^{\alpha + 1}} \,dx\\
    &&+
  \beta \,(K')^\beta\, (1 - p)
  \int_{0}^\infty
  u\big( (1 - \phi) \ex^r + \phi\, \ex^x \big)\,
  {1 + b \1(x \le  q) \over (K' + x)^{\beta + 1}} \,dx 
  - b \,u(\delta v) \,F_X(q) \,.\label{eq:pp}
\eeam
\par
We observe the following property which also applies to \eqref{eq:xxie1.0}. 
\ble\label{thrm:I}
 Assume the two-sided Pareto model \eqref{eq:pareto} and that the utility
 \fct\ $u$ is increasing and differentiable. Then $\frac{\pd \wt
   u_{\rm max}}{\pd \alpha} > 0$ and $\frac{\pd \wt u_{\rm max}}{\pd
   K} < 0$.
\ele
The proof is given in Appendix~\ref{sec:thrmI_proof}. 
\par
We conclude that
$\wt u_{\rm max}$ increases with $\alpha$ and decreases with $K$. Therefore there is a curve of
equal preference on the $(\alpha, K)$-plane. Moving along this
curve in the direction of increasing $\alpha$, one expects the values
of $K$ to increase too, i.e., the estimated values of $\alpha$ and $K$
should appear positively dependent. This is indeed the case for some
real return data, e.g. the ``Energy'', ``Consumer Staples'' and
``Information Technology'' sectors of the S\&P 500 index; see 
Figure~\ref{fig:sectors_parameters}.
\par
We illustrate Lemma~\ref{thrm:I} in Figure~\ref{fig:preference_pareto} for the power-utility
\fct\ \eqref{eq:u}.
For this reason, we replace $\phi$ in the first integral in \eqref{eq:pp} by the optimal $\hat\phi$. The resulting
quantity is denoted by $I(\alpha, K)$.
\par
The value of $\hat\phi$ can be calculated
by numerical integration and optimization with respect to $\phi$ for given values of $K, K', \alpha, \beta$. 
This is shown in 
Figure~\ref{fig:phi_hat_pareto} for the power utility \fct\ \eqref{eq:u}.
\begin{figure}[htb!]
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto5e-1.pdf}    
  \end{minipage}\hfill
  \begin{minipage}{0.5\linewidth}
    \includegraphics[width=\textwidth]{phi_hat_pareto4.pdf}
  \end{minipage}
 \caption{The graphs show $\hat\phi$, the optimal equity allocation
    as a \fct\ of $\alpha$ and $K$ in the two-sided Pareto model
    \eqref{eq:pareto} with $K=K'$, $\alpha=\beta$. We choose
    the utility \fct\ $u$ from \eqref{eq:u}
    for $\xi = 1/2$ (left) and $\xi = 4$
    (right).
  }
  \label{fig:phi_hat_pareto}
\end{figure}
